Creating the ``r params$package_name`` R package

### 2025-07-09 create-FSHybridPLS.Rmd
delete meaningless test

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
```{r classdef:predictor_hybrid}

@@

  # Check functional coefficients (use coef() not $coefs)
  testthat::expect_equal(drop(coef(obj_single$functional_list[[1]])), 1:4)
  testthat::expect_equal(drop(coef(obj_single$functional_list[[2]])), 5:8)
})

@@

###
```{r}
get_CJ <- function(W){
  n <- W$n_sample
  K <- W$n_functional
  ## C_J : C multiplied by J. Scalable to arbitrary K
  predictor_first <- W$functional_list[[1]]
  C_J <- t(predictor_first$coefs) %*% W$gram_list[[1]]
  for (i in 2:K){
    predictor_now <- W$functional_list[[i]]
    C_J <- cbind(
      C_J,
      t(predictor_now$coefs) %*% W$gram_list[[i]]
    )
    }
    return(C_J)
}
```

### get_pre_corrcov
construct the matrix V_star:
$$
    V^\ast = n^{-2}\begin{bmatrix}
         J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z} \\
        \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z}
    \end{bmatrix} = n^{-2} J^* \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^*
    \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$
where $J$ is defined in `get_gram_matrix_block`
equation (7), 
```{r}
#' Compute V matrix (pre-corrected covariance estimator)
#'
#' Constructs the V matrix involving inner products of functional predictors
#' (weighted by their Gram matrices) and scalar predictors against response y.
#'
#' @param W A `predictor_hybrid` object.
#' @param y A numeric response vector of length equal to number of samples.
#'
#' @return A matrix representing the corrected cross-covariance structure.
#' @export
get_pre_corrcov <- function(W, y){
  n <- W$n_sample
  K <- W$n_functional

@@

test code:

| Case | Description               | Functional Coefficients         | Scalar Covariates      | Response         | Expected Behavior                              |
|------|---------------------------|----------------------------------|------------------------|------------------|------------------------------------------------|
| 1    | Constant structure        | All ones or twos                | All ones               | Increasing vector| Output matrix is symmetric and non-negative    |
| 2    | Random structure          | Random normal entries           | Random normal values   | Random vector    | Output is symmetric and numerically PSD        |

```{r}
testthat::test_that("get_pre_corrcov handles constant and random coefficients", {
  suppressPackageStartupMessages(library(fda))

  # Constant coefficients
  fd_const1 <- make_fd(matrix(1, 4, n_sample), basis)
  fd_const2 <- make_fd(matrix(2, 4, n_sample), basis)
  Z_const <- matrix(1, n_sample, n_scalar)
  y_const <- matrix(1:n_sample, ncol = 1)
  obj_const <- predictor_hybrid(Z = Z_const, functional_list = list(fd_const1, fd_const2))

  V_const <- get_pre_corrcov(obj_const, y_const)
  testthat::expect_equal(dim(V_const), rep(sum(obj_const$n_basis_list) + obj_const$n_scalar, 2))
  testthat::expect_true(isSymmetric(V_const))
  testthat::expect_true(all(V_const >= 0))

  # Random coefficients
  set.seed(123)
  fd_rand1 <- make_fd(matrix(rnorm(4 * n_sample), 4, n_sample), basis)
  fd_rand2 <- make_fd(matrix(rnorm(4 * n_sample), 4, n_sample), basis)
  Z_rand <- matrix(rnorm(n_sample * n_scalar), n_sample, n_scalar)
  y_rand <- matrix(rnorm(n_sample), ncol = 1)
  obj_rand <- predictor_hybrid(Z = Z_rand, functional_list = list(fd_rand1, fd_rand2))

  V_rand <- get_pre_corrcov(obj_rand, y_rand)
  testthat::expect_equal(dim(V_rand), rep(sum(obj_rand$n_basis_list) + obj_rand$n_scalar, 2))
  testthat::expect_true(isSymmetric(V_rand))
  testthat::expect_true(all(eigen(V_rand, only.values = TRUE)$values >= -1e-8))
})

  J_star <- get_gram_matrix_block(W)
  alpha <- solve(J_star, t(b))                # unconstrained maximizer
  xi_hat <- alpha / sqrt(t(alpha) %*% J_star %*% alpha)  # normalize

```{r}
  C_J <- get_CJ(W)
  A <- cbind(C_J, W$Z)                      # shape: n × d
  b <- t(y) %*% A / W$n_sample   # equivalent to V_star

  J_star <- get_gram_matrix_block(W)
  alpha <- solve(J_star, t(b))                # unconstrained maximizer
  xi_hat <- alpha / sqrt(sum((alpha) * (J_star %*% alpha) )) # normalize
t(y) %*% A %*% xi_hat
```

```{r}
 V_star <- get_pre_corrcov(W, y)
L <- chol(J_star)
  invL <- Matrix::chol2inv(L)
  A <- V_star %*% t(invL)       # A = V* × t(inv(L))
  E <- invL %*% A               # E = inv(L) × A

  xi_star <- t(invL) %*% e      # xi_star solves t(L) xi = e
```

### get_pls_comp
Computes the first PLS component from a hybrid predictor and response vector using regularization.

@@

```{r}
suppressPackageStartupMessages({
  library(fda)
  library(Matrix)
  library(ggplot2)
})

# --- Helper: Generate realistic smooth functional data ---
generate_fd_data <- function(n_sample, nbasis, freq = 1) {
  basis <- create.bspline.basis(c(0, 1), nbasis = nbasis, norder = 4)
  time_points <- seq(0, 1, length.out = nbasis)
  coefs <- outer(1:nbasis, 1:n_sample, function(i, j) {
    sin(freq * pi * time_points[i]) + rnorm(1, sd = 0.1)
  })
  fd(coef = coefs, basisobj = basis)
}

# --- Settings ---
set.seed(20250604)
n_sample <- 200
n_scalar <- 10
n_functional <- 2
nbasis_vals <- c(4,8,12,16,20)
lambda_vals <- c(0, 1e-8, 1e-6, 1e-4, 1e-2, 1, 10)

results <- data.frame()

# --- Experiment Loop ---
for (nbasis in nbasis_vals) {
  for (lambda in lambda_vals) {
    cat("Trying nbasis =", nbasis, ", lambda =", lambda, "\n")

    # Generate functional predictors
    fd1 <- generate_fd_data(n_sample, nbasis, freq = 1)
    fd2 <- generate_fd_data(n_sample, nbasis, freq = 2)
    Z <- matrix(rnorm(n_sample * n_scalar), n_sample, n_scalar)
    y <- rnorm(n_sample)

    W <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))
    lambda_vec <- rep(lambda, n_functional)

    # Try to compute constraint matrix L
    L <- get_constraint_matrix_hybrid(W, lambda_vec)
    constr_mat_chol <- t(chol(L))  # <-- do not change this

    # Time the PLS computation
    elapsed <- system.time({
      out <- tryCatch({
        pls <- get_pls_comp(W, y, constr_mat_chol)

        if (!is.list(pls) || (is.character(pls) && pls == "stop")) {
          data.frame(
            nbasis = nbasis,
            lambda = lambda,
            eigen_val = NA,
            is_real = FALSE,
            reason = "stop (complex eigenvector)",
            time_sec = NA_real_
          )
        } else {
          data.frame(
            nbasis = nbasis,
            lambda = lambda,
            eigen_val = Re(pls$eigen_val),
            is_real = TRUE,
            reason = "success",
            time_sec = NA_real_  # placeholder, will update below
          )
        }
      }, error = function(e) {
        data.frame(
          nbasis = nbasis,
          lambda = lambda,
          eigen_val = NA,
          is_real = FALSE,
          reason = "error in get_pls_comp",
          time_sec = NA_real_
        )
      })
    })

    out$time_sec <- elapsed["elapsed"]
    results <- rbind(results, out)
  }
}

# --- Print successful configurations ---
cat("\n✅ Successful runs with real eigenvalues:\n")
print(subset(results, is_real == TRUE))

# --- Fastest and slowest ---
cat("\n⏱️ Top 5 fastest settings:\n")
print(head(results[order(results$time_sec), ], 5))

cat("\n🐢 Top 5 slowest settings:\n")
print(tail(results[order(results$time_sec), ], 5))

# --- Visualize success/failure ---
ggplot(results, aes(x = nbasis, y = lambda, fill = reason)) +
  geom_tile(color = "white") +
  scale_y_log10() +
  scale_fill_manual(values = c(
    "success" = "green",
    "stop (complex eigenvector)" = "orange",
    "error in get_constraint_matrix_hybrid" = "red",
    "error in get_pls_comp" = "darkred"
  )) +
  labs(
    title = "get_pls_comp Outcomes: λ vs. # Basis Functions",
    x = "# of Basis Functions",
    y = "Smoothing Parameter λ (log scale)",
    fill = "Result Type"
  )

@@

set.seed(011111)
n_sample <- 100
n_scalar <- 3

# --- Step 1: Zero functional predictors ---
zero_basis <- create.bspline.basis(c(0, 1), nbasis = 4)
zero_fd <- fd(coef = matrix(0, nrow = 4, ncol = n_sample), basisobj = zero_basis)
functional_list <- list(zero_fd, zero_fd)  # two zero-valued functional predictors

# --- Step 2: Simulate scalar predictors and response ---
Z <- matrix(rnorm(n_sample * n_scalar), ncol = n_scalar)
Z <- scale(Z, center = TRUE, scale = FALSE)
W <- predictor_hybrid(Z = Z, functional_list = functional_list)

beta_true <- rnorm(n_scalar)
y <- Z %*% beta_true + rnorm(n_sample, sd = 0.1)
y <- scale(y, center = TRUE, scale = FALSE)

# --- Step 3: Compute constraint matrix ---
lambda <- rep(0, W$n_functional)
L <- chol(get_constraint_matrix_hybrid(W, lambda))

# --- Step 4: PLS with get_pls_comp ---
pls <- get_pls_comp(W, y, L)
xi_hat <- pls$xi_hat

# --- Step 5: PLS via raw Euclidean eigenproblem ---
V_star_vec <- as.vector(t(Z) %*% y)     # equivalent to colSums(Z * y)

J <- crossprod(Z)                       # J = Z^T Z
L_Z <- chol(J)
invL <- solve(L_Z)
E <- invL %*% tcrossprod(V_star_vec) %*% t(invL)

eig <- eigen(E)
e <- eig$vectors[, 1]
xi_manual <- backsolve(L_Z, e, transpose = TRUE)  # xi = L^T \ e

# --- Step 6: get_pls_comp_simple ---
xi_simple <- get_pls_comp_simple(W, y)
xi_simple_vec <- tail(xi_simple$Z, 1)
xi_direct <- get_xi_direct(W, y)
# --- Step 7: Compare all three scalar parts ---

# From get_pls_comp
xi_hat_vec <- tail(xi_hat$Z, 1)

# Cosine similarities
cos_eig_manual <- sum(xi_manual * xi_hat_vec) / (sqrt(sum(xi_manual^2)) * sqrt(sum(xi_hat_vec^2)))
cos_simple_manual <- sum(xi_manual * xi_simple_vec) / (sqrt(sum(xi_manual^2)) * sqrt(sum(xi_simple_vec^2)))
cos_eig_simple <- sum(xi_simple_vec * xi_hat_vec) / (sqrt(sum(xi_simple_vec^2)) * sqrt(sum(xi_hat_vec^2)))

# --- Step 8: Report ---
cat("Cosine similarity (get_pls_comp vs manual eig):", cos_eig_manual, "\n")
cat("Cosine similarity (get_pls_comp_simple vs manual eig):", cos_simple_manual, "\n")
cat("Cosine similarity (get_pls_comp vs get_pls_comp_simple):", cos_eig_simple, "\n")

>>>>

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}

@@

  # Check functional coefficients (use coef() not $coefs)
  testthat::expect_equal(drop(coef(obj_single$functional_list[[1]])), 1:4)
  testthat::expect_equal(drop(coef(obj_single$functional_list[[2]])), 5:8)
})
```

@@

### Spectral method
The old implementation that computes the leading eigenvector of

### helper: get_CJ
```{r}
get_CJ <- function(W){
  n <- W$n_sample
  K <- W$n_functional
  ## C_J : C multiplied by J. Scalable to arbitrary K
  predictor_first <- W$functional_list[[1]]
  C_J <- t(predictor_first$coefs) %*% W$gram_list[[1]]
  for (i in 2:K){
    predictor_now <- W$functional_list[[i]]
    C_J <- cbind(
      C_J,
      t(predictor_now$coefs) %*% W$gram_list[[i]]
    )
    }
    return(C_J)
}
```

### helper: get_pre_corrcov 
construct the matrix V_star:
$$
    V^\ast = n^{-2}\begin{bmatrix}
         J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z} \\
        \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z}
    \end{bmatrix} = n^{-2} J^* \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^*
    \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$
where $J$ is defined in `get_gram_matrix_block`
equation (7), 
```{r}
#' Compute V matrix (pre-corrected covariance estimator)
#'
#' Constructs the V matrix involving inner products of functional predictors
#' (weighted by their Gram matrices) and scalar predictors against response y.
#'
#' @param W A `predictor_hybrid` object.
#' @param y A numeric response vector of length equal to number of samples.
#'
#' @return A matrix representing the corrected cross-covariance structure.
#' @export
get_pre_corrcov <- function(W, y){
  n <- W$n_sample
  K <- W$n_functional

@@

### main function: get_pls_comp
Computes the first PLS component from a hybrid predictor and response vector using regularization.

@@

## linear system method

  J_star <- get_gram_matrix_block(W)
  alpha <- solve(J_star, t(b))                # unconstrained maximizer
  norm_squared <- as.numeric( t(alpha) %*% J_star %*% alpha )
  xi_hat <- alpha / sqrt(norm_squared)  # normalize

@@
